Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-29 paper.qmd
Updated Figure 9

```{r}
#| label: hexbin-regular-two-curvy2
## hexagon binning to have regular hexagons
hb_obj_two_curvy2 <- hex_binning(
  data = umap_two_curvy_scaled, 
  bin1 = 14, 
  r2 = r2)

a1_2 <- calc_bins_y(
  bin1 = 14, 
  r2 = r2
  )$a1

@@

```{r}
#| label: hexbin-regular-two-curvy3
## hexagon binning to have regular hexagons
hb_obj_two_curvy3 <- hex_binning(
  data = umap_two_curvy_scaled, 
  bin1 = 17, 
  r2 = r2)

a1_3 <- calc_bins_y(
  bin1 = 17, 
  r2 = r2
)$a1

@@

## Find the minimum MSE when have duplicate a1
error_two_curvy_clust_umap <- error_two_curvy_clust_umap |>
  group_by(a1) |>
  filter(MSE == min(MSE)) |>
  ungroup() |>
  mutate(prop_dens = b_non_empty/side_length)

@@

mse_two_curvy_b <- ggplot(error_two_curvy_clust_umap, 
                     aes(x = a1, 
                         y = MSE)) +
  geom_vline(xintercept = 0.06,
             linetype="solid",
           color = "#54278f", linewidth=1) +
  geom_vline(xintercept = 0.04,
             color = "#006d2c", linetype=2,
             linewidth=1) +
  geom_vline(xintercept = 0.12, linetype=2,
             color = "#08519c", linewidth=1) +
  geom_line(linewidth = 0.5) + 
  geom_point(size = 1) +
  # scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
  ggtitle("(a)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

prop_dens_a1 <- ggplot(error_two_curvy_clust_umap,
                     aes(x = a1,
                         y = prop_comp)) +
  geom_vline(xintercept = 0.06,
             linetype="solid",
           color = "#54278f", linewidth=1) +
  geom_vline(xintercept = 0.04,
             color = "#006d2c", linetype=2,
             linewidth=1) +
  geom_vline(xintercept = 0.12, linetype=2,
             color = "#08519c", linewidth=1) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) +
  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("proportion of densities")) +
  ggtitle("(d)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

a1_m_two_curvy <- ggplot(error_two_curvy_clust_umap,
                     aes(x = a1,
                         y = prop_bins)) +
  geom_vline(xintercept = 0.12, linetype="solid",
             color = "#006d2c", linewidth=1) +
  geom_vline(xintercept = 0.10, linetype=2,
             color = "#54278f", linewidth=1) +
  geom_vline(xintercept = 0.18, linetype=2,
             color = "#08519c", linewidth=1) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) +
  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
  ggtitle("(b)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

mse_two_curvy_lwd <- ggplot(error_rm_two_curvy, 
                     aes(x = benchmark_rm_lwd, 
                         y = MSE,
                         color = bin1)) + 
  geom_point(
    size = 1
    ) +
  geom_line(
    linewidth = 0.3
    ) + 
  # geom_vline(xintercept = benchmark1, linetype="solid", 
  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
  # scale_x_continuous("standardized bin count", 
  #        transform = "reverse") + 
  scale_x_continuous(breaks = unique(error_rm_two_curvy$benchmark_rm_lwd)) +
  scale_color_manual(values=c("#08519c", "#006d2c", "#54278f")) +
  xlab("threshold to remove low-density hexagons") +
  ylab("MSE") +
  ggtitle("(c)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none")
```

>>>>

```{r}
#| label: hexbin-regular-two-curvy2
## hexagon binning to have regular hexagons
hb_obj_two_curvy2 <- hex_binning(
  data = umap_two_curvy_scaled, 
  bin1 = 13, 
  r2 = r2)

a1_2 <- calc_bins_y(
  bin1 = 13, 
  r2 = r2
  )$a1

@@

```{r}
#| label: hexbin-regular-two-curvy3
## hexagon binning to have regular hexagons
hb_obj_two_curvy3 <- hex_binning(
  data = umap_two_curvy_scaled, 
  bin1 = 23, 
  r2 = r2)

a1_3 <- calc_bins_y(
  bin1 = 23, 
  r2 = r2
)$a1

@@

## Find the minimum MSE when have duplicate a1
error_two_curvy_clust_umap <- error_two_curvy_clust_umap |>
  group_by(a1) |>
  filter(bin1 == min(bin1)) |>
  ungroup() |>
  mutate(prop_dens = b_non_empty/side_length)

@@

mse_two_curvy_b <- ggplot(error_two_curvy_clust_umap, 
                     aes(x = a1, 
                         y = MSE)) +
  geom_vline(xintercept = 0.09,
             linetype="solid",
           color = "#d95f02", linewidth=1) +
  geom_vline(xintercept = 0.06,
             color = "#1b9e77", linetype=2,
             linewidth=1) +
  geom_vline(xintercept = 0.12, linetype=2,
             color = "#1f78b4", linewidth=1) +
  geom_line(linewidth = 0.5) + 
  geom_point(size = 1) +
  # scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
  ggtitle("(a)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

prop_dens_a1 <- ggplot(error_two_curvy_clust_umap,
                     aes(x = a1,
                         y = prop_comp)) +
  geom_vline(xintercept = 0.09,
             linetype="solid",
           color = "#d95f02", linewidth=1) +
  geom_vline(xintercept = 0.06,
             color = "#1b9e77", linetype=2,
             linewidth=1) +
  geom_vline(xintercept = 0.12, linetype=2,
             color = "#1f78b4", linewidth=1) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1) +
  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("proportion of densities")) +
  ggtitle("(d)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

a1_m_two_curvy <- ggplot(error_two_curvy_clust_umap,
                     aes(x = a1,
                         y = prop_bins)) +
  geom_vline(xintercept = 0.09,
             linetype="solid",
           color = "#d95f02", linewidth=1) +
  geom_vline(xintercept = 0.06,
             color = "#1b9e77", linetype=2,
             linewidth=1) +
  geom_vline(xintercept = 0.12, linetype=2,
             color = "#1f78b4", linewidth=1) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1) +
  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
  ggtitle("(b)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

mse_two_curvy_lwd <- ggplot(error_rm_two_curvy, 
                     aes(x = benchmark_rm_lwd, 
                         y = MSE,
                         color = bin1)) + 
  geom_point(
    size = 1
    ) +
  geom_line(
    linewidth = 0.5
    ) + 
  # geom_vline(xintercept = benchmark1, linetype="solid", 
  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
  # scale_x_continuous("standardized bin count", 
  #        transform = "reverse") + 
  scale_x_continuous(breaks = unique(error_rm_two_curvy$benchmark_rm_lwd)) +
  scale_color_manual(values=c("#1f78b4", "#d95f02", "#1b9e77")) +
  xlab("threshold to remove low-density hexagons") +
  ylab("MSE") +
  ggtitle("(c)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none")
```
